Optimizing Abstract Abstract Machines

### 2012-11-14 paper/paper.tex

\subsection{Pre-allocating the store}
Hashes are not the most efficent for raw speed. If we commit to using
sequential natural numbers for variables and expression labels, by the
time we finish parsing the input program, we will know exactly how
many addresses we will need. Thus the store can be represented by a
vector the size of the largest label (call it $\mlab_{max}$) assigned, plus
1. The parser should be assigning new, fresh names to variables
anyway, so this change is a one-liner.

This technique by itself only works in 0-CFA, since there are many
more addresses that can be allocated in 1-CFA and higher. To support
general $k$-CFA in a vector representation, we use a growable vector
and a map from addresses to assigned indices. When assignable indices
run out, we grow the vector. The cost of this map is less than a hash
representation for the whole store since typically there are more
references to an address than allocations of it. This is more
space-efficient than preallocating $(\mlab_{max}+1)^k$ addresses
up-front and addressing by the contour converted to polynomial, since
a majority of combinations of labels will not occur, except in
pathological programs. The all up-front allocation again breaks down
when we add compound data, which introduce new kinds of labels
(e.g. left and right components of a tuple at the contour of
creation).

>>>>

\subsection{Pre-allocating the store}

Internally, the algorithm at this stage uses hash tables to model the store.
%
This is because stores used to be distributed to all states, which
required a compact, dynamic representation.
%
But, such a dynamic structure isn't necessary when we know the
structure of the store in advance: we know all possible entries, and
we know its maximum size.

In a monovariant analysis, the domain of the store is
exactly the set of expressions in the program.
%
If we label each expression with a unique natural, the analysis can
index directly into the store without a hash or a collision.
%
Even for polyvariant analyses, it is possible to compute the maximum
number of addresses and similarly pre-allocate either the spine of the
store or (if memory is no concern) the entire store.
